---
title: The “biblionetwork” package
author: Ariel Zane
date: '2018-03-27'
slug: the-biblionetwork-package
categories:
  - package
  - tutorial
  - dataset
  - network analysis
  - graph
tags:
  - candy
  - package
  - tutorial
  - vignette
  - network
draft: TRUE
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(tidy = FALSE, eval = FALSE)
```

After the high of rstudio::conf 2017, I made a (tiny) list of R-related goals for myself:

```{r echo = FALSE, eval = TRUE}
blogdown::shortcode('tweet', '960859452244099072')
```

Happily, I was able to knock the [second one](https://aczane.netlify.com/2018/02/08/the-first-and-namesake-post-is-it-cake/) out back in February, but the first one has taken more work. 

The `biblionetwork` package has just two functions, currently. They are (so creatively named, you guys):
`make_edgelist` : creates edgelists in dataframe format
`make_nodelist` : creates nodelists in dataframe format

I'm not sure if a package already exists out there that performs these functions, but I couldn't find one when I first started working with co-occurrence lists and networks. I did find a great example from [Francois Briatte](https://www.r-bloggers.com/turning-keywords-into-a-co-occurrence-network/), which I was able to modify for my work. I've since changed the script even further, and benchmarked it to gauge speed: [^1]

```{r echo = FALSE, eval = TRUE}
blogdown::shortcode('tweet', '973903870366814209')
```

Mind you, it's still not very fast (especially for very large networks, ~10K nodes and >= 10^6 edges). But it works!

I'm going to walk through it with the "Candy" dataset from [FiveThirtyEight](https://data.fivethirtyeight.com). This is a dataset of candy types, with a bunch of variables describing the candies; the full description of the data can be found [here](https://github.com/fivethirtyeight/data/tree/master/candy-power-ranking). We're going to rework it a little bit and create a network graph of the candies; candy with similar traits should be closer to each other and linked. I'm only interested in part of the dataset (not the percentages at the end), so I'll just select those columns.

```{r eval = FALSE, echo = TRUE}
library(tidyverse)

# load the data and select the desired columns

dat <- read_csv("candy-data.csv") %>%
  select(competitorname, chocolate,	fruity,	caramel,
         peanutyalmondy, nougat, crispedricewafer, hard, bar,
         pluribus)

# check it out

head(dat)
```

```{r eval = TRUE, message = F, echo = FALSE}
library(tidyverse)

dat <- read_csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/candy-power-ranking/candy-data.csv") %>%
  select(competitorname, chocolate,	fruity,	caramel,
         peanutyalmondy, nougat, crispedricewafer, hard, bar,
         pluribus)

head(dat)

```
The descriptive columns are categorical variables - "1"" indicating "yes," (e.g., a "1" in "chocolate" means it is chocolate or contains chocolate), and "0" for "no." We're going to change this coding a little bit and create text instead. There is probably a more elegant way of doing this, but messing with extracting and looping over colnames is something I don't want to touch with a long stick, so...

```{r eval = TRUE, echo = TRUE}

# create new columns with text instead of binary coding

dat_labels <-
  dat %>%
  mutate(choc_lab = ifelse(chocolate == "1", "chocolate", NA),
         fruit_lab = ifelse(fruity == "1", "fruity", NA),
         car_lab = ifelse(caramel == "1", "caramel", NA),
         pea_lab = ifelse(peanutyalmondy == "1",
                                 "peanutyalmondy", NA),
         nou_lab = ifelse(nougat == "1", "nougat", NA),
         crisp_lab = ifelse(crispedricewafer == "1",
                                   "crispedricewafer", NA),
         hard_lab = ifelse(hard == "1", "hard", NA),
         bar_lab = ifelse(bar == "1", "bar", NA),
         plu_lab = ifelse(pluribus == "1", "pluribus", NA)) 

# and check

head(dat_labels)

```
Now! We're going to reshape this. What we ultimately want is a single descriptor column that has concatenated text. Theoretically, we could use `dplyr::unite()`, but it doesn't handle NAs well, so we'll do it a slightly messier way. We'll start with `dplyr::gather()`.

```{r eval = TRUE, message = F, echo = TRUE}
library(magrittr)

# turn into long format, eliminating rows with NA values

# trim columns from dat_labels

dat_labels %<>%
  select(-c(2:10))

# turn to long format, removing NA rows
dat_long <- gather(dat_labels, key = type, value = descriptor, 2:10, na.rm = TRUE) %>%
  select(-type)

# now create a final column that concatenates all present descriptor text per candy

candy <- aggregate(descriptor ~ competitorname, data = dat_long, paste, collapse = ",")

# and check
head(candy)

```
OKAY. We can finally get to more fun stuff. I mean, I guess that was fun. But I'm itching to share my package!

It's hosted on my [github](https://github.com/aczane/biblionetwork):

```{r eval = TRUE, echo = TRUE}

# install from github
devtools::install_github("aczane/biblionetwork")

# load the library
library(biblionetwork)

```

First, we will want to create the nodelist.
```{r eval = TRUE, echo = TRUE}

```

[^1]:[Alexej's post](http://www.alexejgossmann.com/benchmarking_r/) is extremely helpful for benchmarking, btw.

