---
title: I’m Not a Ninja, but I’ve Got Some Slides!
author: Ariel Zane
date: '2018-03-02'
slug: i-m-not-a-ninja-but-i-ve-got-some-slides
categories:
  - xaringan
  - tutorial
  - Rmd
tags:
  - xaringan
  - tutorial
  - walkthrough
  - Rmd
  - css
  - slides
  - presentation
draft: True
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(tidy = FALSE, echo = FALSE)
```

A recent blog post from [Edwin Thoen]("https://edwinth.github.io/blog-new-things/"), suggesting that we blog about things we just learned, prompted me to write this post, because I think that there is some benefit to seeing how other people learn, and to following examples created by others who are not ninja-level experts. 

I spent yesterday putting together some slides at work, using the [xaringan]("https://github.com/yihui/xaringan") package. I largely followed the examples in [Yihui's slide deck]("https://slides.yihui.name/xaringan/"), the [wiki]("https://github.com/gnab/remark/wiki"), and [Alison's slide deck and code]("https://alison.rbind.io/post/r-ladies-slides/s"). I have *very* limited experience in CSS - I've made a flask app in Python for a project, but in general CSS makes me want to throw things and eat my computer, so...that's where I started from. But at the end of the day, I had a basic slide set with R-generated tables, so I thought maybe I could reproduce that here, for anyone who wants to hit the ground running with very limited prior knowledge.

The documentation is all there, and theoretically you should read it, but I tend to learn best by doing; I'll run through the basics so that I understand where to start, but I find it the most helpful to see the .Rmd and the final output, and learn that way.

1. Download [xaringan]("https://github.com/yihui/xaringan"). I work in RStudio, so my instructions and example will be specific to this method. You can install it with:

```{r}

```


2.

